---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "Aidan Clark"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
library(scales)
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
covid_data_count |> 
  filter(race== "Total") |> ggplot(aes(x = state, y = Deaths)) + geom_col()
```



If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. Write code for a plot and describe what you observe?

I made two plots: one showing the proportion of COVID tests by race in several states and one showing the proportion of the population by race in those states.

````{r}
ggplot(covid_data_long, aes(x = count, y = race)) + geom_col(aes(fill = count_type)) + labs(title = "COVID Data by Race")

```
```{r}
covid_data_long |> 
  filter(count_type == "Tests" & race != "Total") |> ggplot(aes(x = state, y= count)) + geom_col(aes(fill = race), position = "fill") + labs(title = "COVID Tests by Race", fill = "Race", y = "Proportion", x = "State") + scale_y_continuous(labels = scales::percent)
```

```{R}
covid_data_long |>
  filter(race != "Total" & state %in% c("CA", "DE", "IL", "KS", "MO", "NV", "RI", "UT")) |> 
  ggplot(aes(x = state, y = population)) + geom_col(aes(fill= race),  position = "fill") + labs(title = "State Populations by Race", x = "State", y = "Proportion") + scale_y_continuous(labels = scales::percent)
```

2. Are there any conclusions you can draw?
Firstly, it's interesting that we only have testing data for a handful of states. I wonder if most states did not collect testing data, or if our dataset just chose not to include it. I was surprised to see that, in most of the states we have data for, the proportion of tests for each race is relatively equal to the proportion of that race in that state's population. ALthough, one discrepancy I noticed was that in Nevada, the Latinx population had disproportionally fewer tests. There is a note that the Latinx population counts may be inaccurate, so perhaps this is a demonstration of that. Also, it appears that AIAN, which compose a very small proportion of each state's population, compose an even smaller proportion of each state's COVID tests.
3. What were you not able to do due to not having the R knowledge?
I really, really wanted to combine the above two graphs so that the bars for each state were directly adjacent to each other. This would allow for easy comparison. I tried to add two geom_col's, but I couldn't get it to work. I'm wondering whether I need a different geom or maybe there's an argument that I'm missing. I had a lot of fun playing around with different filters and operators, and I'd like to explore those more.
4. What other data would be useful to better understand this data?
I would love to have testing data for the rest of the states, to see if the patterns for this selection of states holds for those. Also, it would be interesting if we had data on whether or not the tests came back positive or negative. Perhaps some people have greater access to COVID tests, so they would be more likely to get tested even if they don't have COVID.
